Formalizing Stack Safety as a Security Property\ifcameraready\else\vspace{-0.75cm}\fi

### 2023-05-16 submission5/core.tex
Consolidate discussion on testing

We use a Coq specification of the RISC-V architecture~\cite{Bourgeat2021AMF},
and extend it with a runtime monitor implementing a stack safety
micro-policy. We choose Coq for our implementation
to ensure that our definitions are well-formed and to leverage the
power of the QuickChick
testing framework~\cite{Pierce:SF4}.

@@

To use QuickChick, we developed random test-case generators that produce
(1) an initial RISC-V machine state, including the program to be executed;
(2)
  an initial policy state, tagging instructions and the stack regions appropriately
  for the micro-policy (see \cref{sec:enforcement})
(3)
  a map from instructions to lists of their operations in the security semantics.

To write such generators we build on the work of Hri\c{t}cu et
al. \cite{TestingNI:ICFP, DBLP:journals/jfp/HritcuLSADHPV16}, which
introduced {\em generation by execution}, a technique that produces
programs that lead to longer executions---and hopefully towards more
interesting behaviors as a result. We extend their technique with an additional
degree of statefulness: when complex calling conventions need to be
observed to avoid early failstops, extra care must be taken to ensure
that % any such
calls are well behaved. For example, immediately
after a call we need to initialize any stack-allocated variables
with an increased probability. However, at the same time, we need to
allow for the possibility of {\em not} performing such an initialization
to allow for potential attack vectors to manifest.

@@

@@

Our properties allow us to identify an enforcement mechanism as
not really stack safe, and to validate a possible fix.

>>>>

We use a Coq specification of the RISC-V architecture~\cite{Bourgeat2021AMF},
and extend it with a runtime monitor implementing a stack safety
micro-policy. We choose Coq for our implementation
to ensure that our definitions are well-formed \rb{, to reason about them,} and to leverage the
power of the QuickChick
testing framework~\cite{Pierce:SF4}.

@@

To use QuickChick, we developed random test-case generators that produce
(1) an initial RISC-V machine state, including the program to be executed;
(2)
  an initial policy state, tagging instructions and the stack regions appropriately
  for the micro-policy (see \cref{sec:enforcement}); and
(3)
  a map from instructions to lists of their operations in the security semantics.
%
\rb{((say a few more words about 3?))}
%
\rb{
That is, our generators need to produce not only a low-level program
that exhibits interesting control-flow patterns, but also the policy
meta-information and code annotations that would normally be produced
by the compiler.
}

To write such generators we build on the work of Hri\c{t}cu et
al. \cite{TestingNI:ICFP, DBLP:journals/jfp/HritcuLSADHPV16}, which
introduced {\em generation by execution}, a technique that produces
programs that lead to longer executions---and hopefully towards more
interesting behaviors as a result.
%
\rb{
Generation by
execution receives as an input a partially instantiated machine state
and attempts to generate an instruction (or a sequence of instructions
such as a blessed sequence) that makes sense locally (e.g., jumps go
to a potentially valid code location and loads read from a
potentially valid stack location). Then we step the machine and repeat
the process until we generate or execute some target number of
instructions, or reach a point where the machine cannot step
any more.
}

We extend their technique with an additional
degree of statefulness:
\rb{
we keep track of the control flow behavior of
the program being generated. Each time a call or return sequence is
generated, we ensure that the appropriate policy tags and code
annotations are set for the entry or return points.
}
When complex calling conventions need to be
observed to avoid early failstops, extra care must be taken to ensure
that % any such
calls are well behaved. For example, immediately
after a call we need to initialize any stack-allocated variables
with an increased probability. However, at the same time, we need to
\rb{allow the generation to sometimes relax those constraints,
introducing potentially ill-formed flows:}
% allow for the possibility of,
for example, {\em not} performing such an initialization
\rb{or \emph{not} returning to the proper call site, so as}
to allow for potential attack vectors to manifest.
\rb{
This causes our programs
to failstop when executed in conjunction with the policy monitor,
but also allows for revealing errors in our setup.}

@@

\rb{
Naturally, testing also revealed a number of errors in our
implementation of the enforcement mechanism (the original was written in C++
and targeted ARM machine code;
%\bcp{right?}\leo{yeah}
we re-implemented it in Coq targeting RISC-V).  These errors range
from trivial typos to ones that require an intriguingly complex setup
to provoke.  The most interesting bug (included in the table as row
{\em HEADER\_NO\_INIT}) was that, on our first try, the blessed call
sequence %/policy combination\apt{??}
did not initialize all locations for the
newly allocated stack frame correctly, but left some of them as
$\tagNoDepth$. This allowed for a potential integrity violation, but
only if a rather complicated sequence of events occured.
The smallest counterexample requires calling a function {\tt f},
which fails to initialize some of its frame during the blessed sequence,
but writes into an uninitialized location $l$ later, treating \(l\) as outside
the stack. Then {\tt f} calls a further function {\tt g} (which should have
the effect of sealing $l$ for integrity purposes). {\tt g} attempts to write to $l$,
which is allowed because the enforcement mechanism still has
$l$ tagged as $\tagNoDepth$, but violates the integrity property on {\tt f}'s data.
%\sna{I believe what went wrong was that we were off-by-one in {\tt main}'s initialization,
%  and the write from {\tt f} was already a violation.}
}

@@

Our properties allow us to identify an enforcement mechanism as
not really stack safe, and to validate a possible fix.
\rb{
((Cache pressure discussion relevant?))
%% Unfortunately, Lazy Per-activation Tagging and Clearing
%% puts severe pressure on the tag cache, so it may not be usable in practice.
%% However, we conjecture that no lazy scheme can be fully stack safe
%% without using unique per-activation identifiers.
}
